### 2016-01-13 00-abstract.tex
Initial commit

% 1. what is the field of your work?
% 2. what is the problem you solve?
% 3. what is your solution/approach?
% 4. what are your results?