### 2017-05-03 gpu-01-intro.tex
Import all texts from IPDPS2016 and ParLearning2016

\section{Introduction}

The past decade has witnessed a tremendous increase in the applicability and
usefulness of artificial intelligence in our daily lives. Much of this activity
was fueled by the mainstream adoption of machine learning approaches as they
provide tools to solve inherently difficult problems.
%
However, many of these techniques require a huge amount of computation
which severely limits the
scale of problems that can be tackled.

In this paper, we present the methodology followed in designing and
implementing a high-performance version of an existing Stochastic Gradient Markov Chain
Monte Carlo (SG-MCMC) machine learning algorithm that detects overlapping
communities in graphs. The algorithm analyzes pair-wise interactions between
entities in order to discover hidden attributes.
%
For instance, consider a social network represented as an undirected graph
where the vertices represent individuals and edges represent relations between
them. Given the relation information, the algorithm can identify latent groups
of individuals that represent shared interests.
%
This problem structure differs from graph partitioning or clustering as there
is a many-to-many relationship between individuals and interests. For example,
each individual can have multiple interests. Simultaneously, each interest group can
span multiple individuals.
%
Formally, this problem domain is known as Mixed-Membership Stochastic
Blockmodels (MMSB). The theory behind the algorithm is discussed in more detail
in~\cite{DBLP:journals/corr/LiAW15}.

The focus of this work is on the computational efficiency and parallel
performance of the SG-MCMC algorithm. More specifically, we discuss the process
of accelerating the algorithm by developing aggressive optimizations targeting
multi-core CPUs and GPUs. The parallel algorithm achieves speedup factors up to 86, compared to a
well-tuned sequential C++ program which itself is a factor~1000-1500 faster than
the original Python/Numpy program developed by the algorithm's authors. From a
computational point of view, this algorithm differs from widespread machine
learning algorithms in several ways. First, it is highly data-intensive which
makes GPU acceleration particularly challenging. Second, owing
to the algorithm's stochastic nature, the majority of its memory access
patterns and data dependencies are non-deterministic. As a result,
straightforward
optimization attempts of the memory access patterns either fail or lead
to non-intuitive results.

Through careful analysis of the computation and data structures we show that
the algorithm's full state can be reduced by roughly 75\%. Compressing the
state significantly reduces the data intensity and allows for tackling larger
problems while maintaining all state in memory.
%
Further, by cataloguing and accounting for the various load and store
operations, we identified the highest priority locations of data reuse. In
order to circumvent the unclear optimization landscape, we developed an
effective kernel code generation mechanism that explores the benefits of
exploiting all permutations of the available optimization opportunities. These
optimizations include caching in shared memory, caching in the register file,
loop unrolling and explicit vectorization.

In summary, the contributions of this
work are:
\begin{itemize}
  \item Decreasing the algorithm's data intensity by eliminating 75\% of its
    memory footprint
  \item Tuning the algorithm's performance by maximizing data reuse and
    identifying the highest performing combination of optimizations through
    dynamic kernel code generation
  \item Achieving speedup factors of 21 and 86 over an optimized sequential
    program using a multi-core CPU and a GPU respectively
  \item A comparative performance analysis of the accelerated algorithm
    versions on a multi-core CPU and four GPUs, highlighting the particular
    optimization combinations that were successful per device.
\end{itemize}

The remainder of this paper is organized as follows. A description of the
sequential version of the algorithm and its data structures is provided in
Section~2. Section~3 delves into the design of the parallel algorithm and
discusses its contributions.  Section~4 provides an empirical evaluation of the
contributions of this work.  Section~5 presents an overview of related works.
Finally, Section~6 concludes.