### 2017-05-03 gpu-04-04-distributed.tex

\begin{comment}
3 subsections:
   1. intro and motivation
   2. design
   3. evaluation

1. intro and motivation
     if K * N exceeds single-host memory need solution: use distributed memory, by
     partitioning pi/phi
     expect things to be memory-access dominated, so:
      - because of speed, cannot use disk or even SSD; fully random accesses, no data
        reuse or locality
      - need fastest remote memory available; introduce RDMA
     trade speed for memory in many places...

2. design
   - use master/worker paradigm

   - memory allocations:
     + distributed key-value store for pi/phi
       choose RDMA read/write; no synch, no concurrency, fixed size for reads
       as well as writes
     + distribute pi/phi over the workers
     + keep the graph at the master
     + distribute the held-out set over the master + workers

   - implementation:
     iterate:
     + master draws mini-batches
     + master scatters mini-batch and graph slice over the W workers
     + workers load pi
     + workers draw neighbors (OpenMP)
     + workers calculate updated phi (m/W * n) (OpenMP)
     + barrier
     + workers calculate and store updated pi (OpenMP)
     + barrier
     + master calculates beta (m) (OpenMP)
     + if perplexity iteration:
       barrier
       master and workers calculate perplexity, summed w/ reduce (OpenMP/MPI)

   - details
     + use google sparseset because std::unordered_set uses 32 bytes of overhead
       per item
     + have a look at RDMA bandwidth as a function of N and W

\end{comment}

\subsection{Distributed-memory solution}

The previous subsection described how an optimized MCMC a-MMSB solver
is implemented, but that implementation is limited to the memory of one
GPU or CPU. This means that the largest problems in the SNAP network
overview cannot be tackled in this way. The network graph requires to
store at least $N$ nodes and $E$ edges, and this requires at least $N +
2E$ 4-byte integers. The storage required for the~$\pi$ field can easily
exceed this: $N \cdot K$ 8-byte doubles. For example, to process the
Friendster graph for $K=4096$ communities, the memory for~$\pi$ alone
requires $N \cdot K$ doubles, i.e.\ 2TB of storage, which far exceeds
the memory of most current servers. Note that computing time can be
traded for memory because only one of the~$\pi$ and~$\phi$ fields needs
to be stored in full; the stored one can be used to recompute the other,
which is a fast, trivially parallel operation.

There are two important observations for a solution to this problem. First,
this type of solver performs a limited amount of computation per data item
that is loaded; therefore it is imperative to provide the fastest possible
access to~$\phi/\pi$ rows that are stored outside main memory. Second,
there is little or no locality in the rows of~$\phi/\pi$ that are accessed,
so opportunities for caching data accesses are very small. Consider the
dominant \texttt{update\_phi} calculation, which requires a row of~$\phi$
for each of the mini-batch members and $n$ rows of neighbor~$\phi$ per
mini-batch node. Only the mini-batch node~$\phi$ accesses can be cached, $1/n$
of all accesses.

In any case, fulfilling the memory requirements will be time-costly. Out
of hand, we discarded the option to use disks; disks, even SSDs, are
orders of magnitude slower than local memory access, and because of the
nature of the algorithm, execution speed will suffer proportionally.

The solution we investigate in this paper is to use the aggregate
memory of a cluster of machines by partitioning~$\pi$ over the memory of
the cluster machines. Our implementation uses Infiniband RDMA (Remote
Direct Memory Access)~\cite{RDMA}, a technology that allows a machine to
read or write the memory of another machine through their NICs (network
interface cards). The fastest RDMA primitives do not even involve the
remote machine host {\em at all}; all memory transfers are performed by
the remote NIC through DMA operations.

\subsection{The RDMA distributed key-value store}

The distributed storage of~$\pi$ is implemented as a D-KV (distributed
key-value) store: the key is the index, the value is a row of~$\pi$. The
key-value store is narrowly tailored to the distributed solver, because
that allows a very fast implementation:
\begin{itemize}
\item the values are fixed-size: an array of $K$ doubles
\item the layout of the data is static: at startup, the key-value store is
filled with~$\pi[node]$ values; during the algorithm's lifetime, there are
only reads and overwrites, no deletes or new inserts
\item the keys are contiguous in the range $0..N-1$, so the value address
can be calculated trivially from the key
\item the computation is separated into phases of reading~$\pi$ and phases
of writing~$\pi$: there are no concurrent write or read/write accesses, so
there is no need for concurrency control at all
\end{itemize}
These properties allow our D-KV store to use only RDMA reads and RDMA writes,
which means that there is never any involvement from the remote host.

\subsection{Design of the distributed algorithm}

The distributed algorithm uses MVAPICH2\cite{mvapich2} for all communication
except the RDMA D-KV store, and OpenMP\cite{openmp} for thread parallelism. The
implementation follows a master/slave paradigm. The master stores the full
network graph~$G$, the workers each store a partition of~$\pi$. \textit{ASIDE:
We decided not to scatter the graph over the workers, so all their memory can
be used to store~$\pi$.} In each iteration, the master draws the mini-batch
and partitions it evenly over the workers. The master also sends the slices
of $G$ accessed by each part of the mini-batch to its owner worker. The
workers independently draw a neighbor set for each mini-batch node that they
own. Then each worker loads~$\pi$ from the RDMA key-value store for its
mini-batch and neighbors nodes. Each worker then independently calculates,
with threads in parallel, an updated value for~$\phi$ for its mini-batch
nodes.  A barrier before the update to~$\pi$ through the D-KV store ensures
consistency. The master then calculates the update to $\beta/\theta$,
using thread parallelism; the amount of work in \texttt{update\_beta} is
too little to warrant a distributed implementation. Then, if applicable,
a perplexity calculation is performed in parallel by all machines in the
cluster. Each machine has its partition of the held-out set. It calculates
its contribution, with thread parallelism, to the perplexity value, and those
contributions are summed through an MPI reduce operation.

As we pointed out in Section~\ref{sec-algorithm}, the computational complexity for
\textbf{update\_phi} is $m \cdot n \cdot K$, with $m$ the number of nodes in the
mini-batch and $n$ the size of the neighbor sampler, which is independent of the size
of $G$. The perplexity calculation has complexity $h \cdot K$, with $h$ the size of
the held-out set. This is a fraction of the edges of $G$: 1\%~for huge network graphs,
otherwise~10\%.
The computation in our algorithm lends itself well for a distributed
implementation, but the remote $\pi$ accesses will shift the balance between
memory accesses and computation significantly. We will report quantitatively
in our evaluation Section~\ref{section-evaluation-distributed}.