### 2017-05-04 gpu-04-design.tex

\subsection{System Design and Implementation}

The process of designing an accelerated version of the SG-MCMC MMSB algorithm
involved multiple transformations. First, an efficient C++ baseline version
was created. Further, changes to the algorithm and the data structures
were carried out, for both efficient resource utilization and parallelization.
This section provides an overview of the system's evolution in incremental
phases, identifying the key contributions and differences between consecutive
states.

\input{gpu-04-02-sequential}

\input{gpu-04-03-accelerator}