### 2017-05-18 gpu-04-design.tex

% \subsection{System Design and Implementation}

Based on the efficient C++ baseline version,
changes to the algorithm and the data structures
were carried out, for both efficient resource utilization and parallelization.
This section provides an overview of the system's evolution in incremental
phases, identifying the key contributions and differences between consecutive
states. This section concludes with an in-depth evaluation of the performance
of our various optimizations, on multi-core CPUs as well as a number of GPU
accelerators.

% \input{gpu-04-02-sequential}

\input{gpu-04-03-accelerator}