### 2017-05-18 gpu-05-evaluation.tex

\subsection{Evaluation}
\label{sec-evaluation}

The following sections provide a detailed experimental performance analysis
of the optimizations described above. We start by discussing the performance
effects of parallelizing the computations
on a multi-core CPU using OpenCL.
%
Next, we assess the trade-offs associated with each of the GPU optimization
strategies and their performance effects on four GPUs spanning two chip
architecture generations.
%
Finally, we present the overall outcome of accelerating the algorithm by
displaying its fast convergence over the largest networks that fit in the
memory of one GPU or CPU.

As noted in Section~\ref{sec-experimental}, all experiments were conducted on the VU Amsterdam DAS5 cluster.
Table~\ref{table-gpus} presents the properties of the various GPUs in
the DAS5 cluster.

% \input{gpu-05-02-sequential}
\input{gpu-05-03-accelerator}
\input{gpu-05-05-perplexity}