### 2017-05-23 gpu-05-evaluation.tex
Move exp. setup into GPU section

\subsection{Evaluation}
\label{sec-evaluation}

The following sections provide a detailed experimental performance analysis
of the optimizations described above. We start by discussing the performance
effects of parallelizing the computations
on a multi-core CPU using OpenCL.
%
Next, we assess the trade-offs associated with each of the GPU optimization
strategies and their performance effects on four GPUs spanning two chip
architecture generations.
%
Finally, we present the overall outcome of accelerating the algorithm by
displaying its fast convergence over the largest networks that fit in the
memory of one GPU or CPU.

\input{04-experimental-setup}

% As noted in Section~\ref{sec-experimental}, all experiments were conducted on the VU Amsterdam DAS5 cluster.
% Table~\ref{table-gpus} presents the properties of the various GPUs in
% the DAS5 cluster.

% \input{gpu-05-02-sequential}
\input{gpu-05-03-accelerator}
\input{gpu-05-05-perplexity}