### 2017-05-24 gpu-04-design.tex

% \subsection{System Design and Implementation}

Based on the efficient C++ baseline version, a version for GPU accelaration
was constructed.
This involved optimizations for both efficient resource utilization and
parallelization.
This section provides an overview of the system's evolution in incremental
phases, identifying the key contributions and differences between consecutive
states. This section concludes with an in-depth evaluation of the performance
of our various optimizations on a number of GPU
accelerators.

% \input{gpu-04-02-sequential}

\input{gpu-04-03-accelerator}