### 2017-05-29 gpu-04-design.tex
Remove the history approach from OpenCL. Remove OpenCL

% \subsection{System Design and Implementation}

The first contribution of this paper is construction of a version for GPU
acceleration, based on the efficient C++ baseline version.
This involved optimizations for both efficient resource utilization and
parallelization.
This section provides an overview of the system's evolution in incremental
phases, identifying the key contributions and differences between consecutive
states. This section concludes with an in-depth evaluation of the performance
of our various optimizations on a number of GPU
accelerators.

% \input{gpu-04-02-sequential}

\input{gpu-04-03-accelerator}