### 2017-05-31 04-experimental-setup.tex
Reorder tables both for GPUs and SNAP sets

\subsection{Experimental setup}
\label{sec-experimental}

Unless indicated otherwise, the experiments described in the next sections
were conducted on the VU~Amsterdam and Leiden DAS5 clusters.  The clusters
consists of 68 and 24 compute nodes respectively. Each node is equipped with
a dual 8-core Intel Xeon E5-2630v3 CPU clocked at 2.40GHz, 64GB of memory and
8TB of storage. The nodes of each site are interconnected by FDR~InfiniBand.
Additionally, the VU cluster is fitted with a number of Nvidia GPUs including
GTX~\mbox{Titan-X/Pascal},
GTX~\mbox{Titan-X/Maxwell}, GTX980, K40c and K20m.

To assess the performance of our implementations, we use publicly available
community networks from the SNAP collection~\cite{snapnets}, including the largest
networks. Table~\ref{table-snap}
outlines the graph properties of the networks used.

\begin{table}
  \centering
  \def\tabcolsep{0.2em}
  \begin{tabular}{l r r r p{9em}}
    Name            & \#Vertices &       \#Edges & \multicolumn{1}{c}{\#Ground-} & Description \\
                    &            &               & \multicolumn{1}{c}{truths}    &             \\
    \hline
    CA-HepPh        &    12,008  &    118,521    & \multicolumn{1}{c}{n/a}       & Paper citation network \\
    com-DBLP        &    317,080 &     1,049,866 & 13,477         & Computer science bibliography network \\
    com-Amazon      &    334,863 &       925,872 & 75,149         & Product copurchasing network \\
    com-Youtube     &  1,134,890 &     2,987,624 & 8,385          & Video-sharing social network \\
    com-Orkut       &  3,072,441 &   117,185,083 & 6,288,363      & Online social network \\
    com-LiveJournal &  3,997,962 &    34,681,189 & 287,512        & Online blogging social network \\
    com-Friendster  & 65,608,366 & 1,806,067,135 & 957,154        & Online gaming social network \\
    \hline
    \\[-1ex]
  \end{tabular}
  \caption{Summary of SNAP graph data sets used for evaluation.}
  \label{table-snap}
\end{table}