### 2017-06-01 gpu-04-design.tex
Rephrase GPU design intro for better overview

% \subsection{System Design and Implementation}

As the first contribution of this paper, we created optimized versions for
accelerators of the MCMC-aMMSB
algorithm, based on the efficient C++ baseline version.
The algorithm is data-intensive and nondeterministic in its memory accesses,
which makes an efficient GPU implementation highly nontrivial. Hence, profound
changes to the algorithm and the data structures were necessary,
both for efficient resource utilization and parallelization.
% This section provides an overview of the system's evolution in incremental
% phases, identifying the key contributions and differences between consecutive
% states.
This section describes the various optimization steps in technical detail.
It concludes with an in-depth evaluation of the performance
of our optimizations,
on multi-core CPUs as well as a number of GPU accelerators.

% \input{gpu-04-02-sequential}

\input{gpu-04-03-accelerator}