### 2017-06-01 gpu-05-evaluation.tex

\subsection{Evaluation}
\label{sec-evaluation}

The following sections provide a detailed experimental performance analysis
of the optimizations described above.
%
We assess the trade-offs associated with each of the GPU optimization
strategies and their performance effects on five GPUs spanning three chip
architecture generations.
%
Finally, we present the overall outcome of accelerating the algorithm by
displaying its convergence speed over the largest sensible combination of
community network and number of communities while keeping the algorithm data
in the memory of a GPU.

\input{04-experimental-setup}

% As noted in Section~\ref{sec-experimental}, all experiments were conducted on the VU Amsterdam DAS5 cluster.
% Table~\ref{table-gpus} presents the properties of the various GPUs in
% the DAS5 cluster.

% \input{gpu-05-02-sequential}
\input{gpu-05-03-accelerator}
\input{gpu-05-05-perplexity}