### 2017-06-13 01-intro.tex
Process Henri's comments 1june (except Maxwell/Pascal analysis for GPU strategies)

\section{Introduction}
The tremendous amount of data that we generate through our daily applications such as
social networking services, online shopping, and news recommendations, provides us
with an opportunity to extract hidden but useful, even invaluable information.
Realizing this opportunity, however, requires a significant amount of effort because
traditional machine learning algorithms may become inefficient with large amounts of data.

There have been two main approaches to resolving this issue; machine learning researchers have developed new scalable algorithms~\cite{bottou2010large, boyd2011distributed}, while systems and networking researchers have worked on developing new generic software infrastructure systems which can be leveraged to construct machine learning solutions more efficiently~\cite{dean2008mapreduce, chang2008bigtable}.
However, the best possible performance is often achieved by carefully integrating both of these approaches in a single solution.

One such big data problem is analyzing large graphs such as social networks where it is not unusual to see a network consisting of billions of edges and tens of million of vertices~\cite{yang2015defining}. In particular, we are interested in the overlapping community detection problem~\cite{xie2013overlapping}, where the goal is to learn the probability distribution of each vertex to participate in each community, given a set of vertices, the links between them (which are usually very sparse), and the number of latent communities. A community can be seen as a densely connected group of vertices that are only sparsely connected to the rest of the network. This problem is significantly more complex than the related domain of detecting disjoint communities.

The problem of detecting overlapping communities is modeled by the mixed membership stochastic block model (MMSB)~\cite{airoldi2009mixed} and in this paper we are particularly interested in a variant of MMSB, called assortative MMSB (a-MMSB\footnote{Although we work on a-MMSB for simplicity, it is also straightforward to apply the proposed method to the general MMSB model.}) \cite{gopalan2012scalable}.
The MMSB model is a probabilistic graphical model~\cite{koller2009probabilistic} that represents a convenient paradigm for modeling complex relationships between a potentially large number of random variables. Bayesian graphical models, where we define priors and infer posteriors over parameters, also allow us to quantify model uncertainty and facilitate model selection and averaging. However, an increasingly urgent question is whether these models and their inference procedures will be up to the challenge of handling very large graphs.

There have been two main recent advances in this direction of scalable Bayesian inference: methods based on stochastic variational Bayes (SVB)~\cite{gopalan2012scalable,hoffman2013stochastic,gopalan2013efficient} and stochastic gradient Markov chain Monte Carlo (SG-MCMC)~\cite{welling2011bayesian,patterson2013stochastic,ahn2014distributed,ahn2012bayesian}. Both methods have the important property that they only require a small subset of the data for every iteration. In other words, they can be applied to (infinite) streaming data.

In this paper, we focus on the SG-MCMC method applied to the a-MMSB model. Recent
work of the authors of this paper~\cite{LiAW15} showed that this methodology is faster and more accurate than the
SVB method.

From a computational point of view, our SG-MCMC a-MMSB algorithm differs
from widespread machine learning algorithms like deep learning in several
ways. First, it is \emph{highly data-intensive} which makes parallel acceleration
particularly challenging. Second, owing to the algorithm's stochastic
nature, the majority of its memory access patterns and data dependencies
are \emph{non-deterministic}. As a result, straightforward optimization attempts
of the memory access patterns either fail or lead to non-intuitive results. Third,
the size of its intermediate data structures makes it difficult to scale to very large
community graphs.

Starting from an existing implementation in Python/Numpy, we create an efficient
C++ version to serve as the baseline for the parallel versions.
This already improves performance by a factor of 1000 to 1500 compared to
Python.
% This transformation is not considered one of our key contributions.
Then, our key contributions to the algorithm are:

\begin{itemize}
\item \textbf{scaling up to run on GPUs.}
The application is not a natural fit for GPUs, because it is data-bound and
irregular in its data accesses. Nevertheless, we developed aggressive
memory allocation optimizations to solve the data bandwidth issues for GPUs.
This implementation achieves up to 120$\times$ speedup compared to the C++ baseline.
%
We reduced the algorithm's memory requirements by roughly 75\%, which
significantly reduces the data intensity and allows for tackling larger
problems; % while maintaining all state in memory.

\item \textbf{scaling out to overcome the memory bounds} posed by 
single-machine systems, to be the first to solve the largest publicly available
community graphs. % This necessitated overcoming several challenges.
% The algorithm's state grows rapidly with larger graphs and number of latent
% communities.
Since the full state for the larger graphs does not fit in a
single machine's memory, it is partitioned across a cluster
of machines. Hence, the cluster nodes must read
remote memory hosted by their peers. Since the algorithm is data-bound, we
substantially optimized remote memory accesses by implementing
a custom RDMA store.
On top of that, we pipelined the algorithm stages by fetching
data in advance over the network.
\begin{comment} Finally, the algorithm's computation
is effectively distributed across the cluster nodes and parallelized further
within each node by exploiting their multi-core CPUs.
\end{comment}

The use of pipelining makes it pointless to combine GPU acceleration
into the distributed implementation: computing for an iteration takes less
time than fetching the data from remote memory, and the computation is
fully overlapped with fulfilling the data requests for the next iteration.
% Hence, it is useless to improve the computation time.

\end{itemize}
% To this end, we propose a design of a parallel and distributed system specifically tailored to solve the a-MMSB problem. In particular, we use a mixture of OpenMP, MPI and RDMA in order to efficiently scale and accelerate the algorithm's computation.

\begin{comment}
Further, by cataloguing and accounting for the various load and store
operations, we identified the highest priority locations of data reuse. In
order to circumvent the unclear optimization landscape, we developed an
effective kernel code generation mechanism that explores the benefits of
exploiting all permutations of the available optimization opportunities. These
optimizations include caching in shared memory, caching in the register file,
loop unrolling and explicit vectorization.
\end{comment}

A bird's-eye view of the work in this paper, without any technical detail, has been
published at CCGrid-2016~\cite{10.1109/CCGrid.2016.98}.
The results of our distributed MCMC-aMMSB implementation
have been previously published at the 2016 ParLearning
workshop~\cite{DBLP:conf/ipps/El-HelwHLAWB16}, where it earned a best-paper
award.

The remainder of this paper is organized as follows.
Section~\ref{sec-background} provides an overview of the
algorithm and its theoretical foundation, with related work in
Section~\ref{sec-related}. Section~\ref{sec-design-sequential} describes how we
arrived at an efficient sequential version that serves as the baseline for the
parallel implementations. Section~\ref{sec-gpu} delves
into the parallelization optimizations for GPUs.
Section~\ref{sec-distr} describes the design, implementation and evaluation of
the scalable distributed solution. Finally, Section~\ref{sec-conclusion}
provides concluding remarks.

% -------\\
% Community detection is the central problem in network analysis, with the goal of identifying the groups of related nodes that are densely connected within this group but sparsely connected to the rest of the network. Different from classical community detection problem where we assume each node belongs to one single community, our paper considers overlapping communities where each of the nodes might belong to multiple communities. In particular, we consider the model called a-MMSB(Assortive Mixed Membership Stochastic Blockmodel) in this paper, which was first introduced in~\cite{gopalan2012scalable}.\\

% a-MMSB, as a probabilistic graphical model, represents a convenient paradigm for modeling complex relationships between a potentially large number of random variables. It also uses priors and posteriors to quantify model uncertainty and facilitate model selection and averaging. While a-MMSB provides rich representation power, like other Bayesian models, the inference procedures of handling big data which is common in real world is still a very challenging problem. \\

% Consider the large networks such as a social network, it easily runs into billions of edges and tens of million of nodes. In addition to that, the number of communities might exceed few millions. 
% There were two types of scalable algorithm for a-MMSB have been introduced recently \cite{gopalan2012scalable}, stochastic variational Bayesian inference (SVB) and stochastic gradient Markov chain Monte Carlo(SG-MCMC), respectively. Both methods have the important property that each iteration only relies on a small subset of the data. Although both methods can work for some large networks with many hundreds of nodes, the performance is far less satisfactory when it applies to the so-called "big data" such as facebook network with millions of nodes and billions of edges. Given the facts that \cite{LiAW15}, SG-MCMC runs faster and converges to the better local minima, in this paper, we mainly consider the problem of scaling up SG-MCMC to the big data set.  \textit{As far as we know, this is the first paper that studies community detection on the frencter dataset with few billion's of edges.}